---
title: "CM_shotgun_Entamoeba_RFC_Rmd"
author: "MAR"
date: "4/10/2019"
output: html_document
---
---
title: "CM_shotgun_lactose_RFC.Rmd"
author: "MAR"
date: "4/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  dev='svg'
)
```

```{r Libraries, include=FALSE}
library("tidyverse")
library("magrittr")
library("taxonomizr")
library("reshape2")
library("ggplot2")
library("eclectic")
library("pander")
library("vegan")
library("ape")
library("scales")
library("randomForest")
library("plyr")
library("caret")
library("rfUtilities")
library("ggplot2")
library("ggbeeswarm")
library("tidyr")
library("readr")
library("qiimer")
library("doParallel")
library("e1071")
```

This is an R Markdown that summarizes k-mer based classification of shotgun metagenomic reads. Specifically, shotgun reads from Cameroonian fecal samples were analyzed by [Kraken Uniq](https://github.com/fbreitwieser/krakenuniq) using the database of bacterial, archaeal, viral, protozoan, fungal, and helminth sequences downloaded from NCBI nr/nt collection by Arwa in December 2018. This workflow was done using [sunbeam](https://github.com/sunbeam-labs/sunbeam) using the [Kraken Uniq extension](https://github.com/ArwaAbbas/sbx_kraken_uniq).

The output is tab-delimited, with one line per taxon. The fields of the output,from left-to-right,are as follows:  

Percentage of reads covered by the clade rooted at this taxon  
Number of reads covered by the clade rooted at this taxon  
Number of reads assigned directly to this taxon  
Rank code reporting all 27 ranks defined in the NCBI taxonomy.
NCBI taxonomy ID  
Indented scientific name

The output file of KrakenUniq report file is identical to Kraken. The report file has a couple of modifications - namely a header and three additional columns:  
kmers: number of unique k-mers  
dup: average number of times each unique k-mer has been seen  
cov: coverage of the k-mers of the clade in the database

#Functions and themes 
```{r Read Kraken files}
read_kraken_files <- function(filename) {
                              x = read.delim(filename, comment.char="#",
                                             stringsAsFactors = FALSE)
                              samplename = gsub("-reports.tsv", "",filename)
                              x$V10 = samplename
                              return(x)
}
```

```{r Add noise/generate pos control in RFC data}
#A standard deviation of 1 will give you a fairly correlated feature, s.d. of 0.3 will give you a highly correlated feature, and s.d. of 3 will give you an uncorrelated feature.

add_noise <- function (x, sd=1) {
  as.numeric(x) + rnorm(length(x), sd = sd)
}
```
```{r Remove features that are non-zero less than a specified proportion of the time}

remove_rare <- function( table , cutoff_pro ) {
  row2keep <- c()
  cutoff <- ceiling( cutoff_pro * ncol(table) )  
  for ( i in 1:nrow(table) ) {
    row_nonzero <- length( which( table[ i , ]  > 0 ) ) 
    if ( row_nonzero > cutoff ) {
      row2keep <- c( row2keep , i)
    }
  }
  return( table [ row2keep , , drop=F ])
}
```

```{r Plotting}
#Boxplots
theme_aa_boxplot <- theme_bw(base_size = 12) +
  theme(panel.grid.major = element_blank(),
        strip.background.x = element_rect(fill = "white", color = "black"),
        axis.text = element_text(size = rel(1.1), color = "black"),
        axis.title = element_text(size = rel(1.1)),
        axis.ticks.x = element_blank())
#Heatmap
theme_aa_heatmap <- theme_bw(base_size = 12) +
  theme(panel.grid.major = element_blank(),
        strip.background.x = element_rect(fill = "white", color = "black"),
        strip.background.y = element_blank(),
        axis.ticks = element_blank(),
        axis.text.x = element_blank())
#Barchart
theme_aa_bars <- theme_bw(base_size = 12) +
  theme(panel.grid.major = element_blank(),
        strip.background = element_rect(fill = "white", color = "black"),
        axis.text.y = element_text(color = "black"),
        axis.text.x = element_blank(),
        axis.title = element_text(size = rel(1.1)),
        axis.ticks.x = element_blank())

#Scatterplot
theme_aa_scatter <- theme_bw(base_size = 12) +
  theme(panel.grid.minor = element_blank(),
        strip.background = element_rect(fill = "white", color = "black"),
        axis.text = element_text(size = rel(1.1), color = "black"),
        axis.title = element_text(size = rel(1.1)))

#Kind of crazy, but these need to be in the exact same order as they would be as factors in the dataframe.
ethnic_cols <-c("Bagyeli" = "#a9315c",
                "Baka" = "#f58aa5",
                "Bantu" = "#ffcc99", 
                "Mbororo_Fulani" = "#999FFF")
ANTS_shapes <- c("Neg" = 2,
                 "Pos" = 17)

```

#Import data
```{r Set up Taxonomizr, message=FALSE, include=FALSE, cache = TRUE}
#Set up Taxonomizer databases from a shared site if it doesn't already exist.
#This takes a while

if(!exists("taxaNodes")) 
   taxaNodes<-read.nodes('/media/lorax/projects/virome-dark-matter/software/taxonomizr/nodes.dmp') else print("Taxa Nodes already loaded")

if(!exists("taxaNames")) 
  taxaNames<-read.names('/media/lorax/projects/virome-dark-matter/software/taxonomizr/names.dmp') else print("Taxa Names already loaded")

#Test that it works.
taxaId<-accessionToTaxa(c("LN847353.1","AL079352.3"),"/media/lorax/projects/virome-dark-matter/software/taxonomizr/accessionTaxa.sql")
print(taxaId)
taxa_df <- as.data.frame(getTaxonomy(taxaId,taxaNodes,taxaNames))
taxa_df$taxa_id <- rownames(taxa_df)

#Clean up
rm(taxa_df)
rm(taxaId)

```

```{r Kraken annotation}
#Cameroon files
kraken_files <- list.files(path = "/media/THING2/louis/15_RubelCameroonShotgun/01_CameroonShotgun/sunbeam_output/classify/kraken_uniq/report",
                           pattern = glob2rx("*.tsv"),
                           full.names = TRUE)

#Exclude the empty files. Feel like there should be a more elegant way to avoid these.
kraken_files <- kraken_files[grep(pattern = "D0023-reports.tsv", kraken_files, 
                                  invert = TRUE)]
kraken_files <- kraken_files[grep(pattern = "D0041-reports.tsv", kraken_files, 
                                  invert = TRUE)]

kraken_dfs <- lapply(kraken_files, read_kraken_files)

classify_df <- do.call(rbind, kraken_dfs)

colnames(classify_df) <-c("PercentReads", 
                          "CladeNumberReads",
                          "TaxonNumberReads",
                          "Kmers","Dup", "Coverage", 
                          "TaxID", "Rank",
                          "Name", "Sample")

classify_df$Sample <- gsub("/media/THING2/louis/15_RubelCameroonShotgun/01_CameroonShotgun/sunbeam_output/classify/kraken_uniq/report/", 
                           "", 
                           classify_df$Sample)

#Clean up
rm(kraken_files)
rm(kraken_dfs)

```

```{r Metadata}
#Only want the shotgun samples
meta_df <- read.csv("/media/THING2/louis/15_RubelCameroonShotgun/consign-kegg-r-ator/shotgun_metadata_cameroon_5119.csv", stringsAsFactors=FALSE) %$%
  dplyr::mutate(., sampleID = gsub(pattern = "Shotgun_", replacement = "", sampleID)) %>%
  dplyr::filter(sampleID %in% unique(classify_df$Sample)) %>%
  droplevels.data.frame()

  #dplyr::mutate(., sampleID = gsub(pattern = "Shotgun_", replacement = "", sampleID)) %>%
  #dplyr::filter(sampleID %in% unique(classify_df$Sample)) %>%
  #droplevels.data.frame()


rownames(meta_df) = meta_df$sampleID
meta_df = as.data.frame(meta_df[,-1])

meta_df = as.data.frame(t(meta_df))

IDdrops <- c("EB119","EB3","EB75","LibNeg1","LibNeg2")

metadata_IDs_filtered <- meta_df[,!colnames(meta_df) %in% IDdrops]

metadata_IDs_filtered = as.data.frame(t(metadata_IDs_filtered))

newIDdrops <- which(is.na(metadata_IDs_filtered$BothCommensalPathogen))#check for NAs, note for downstream steps
newIDdrops <- names(newIDdrops)
metadata_IDs_filtered = as.data.frame(t(metadata_IDs_filtered))

metadata_IDs_filtered <- metadata_IDs_filtered[,!colnames(metadata_IDs_filtered) %in% newIDdrops] 
metadata_IDs_filtered = as.data.frame(t(metadata_IDs_filtered))


```

```{r Find the missing samples, eval=FALSE, include=FALSE}
#Find the samples that have been analyzed by shotgun sequencing but are not in metadata
setdiff(classify_df$Sample, metadata_IDs_filtered$sampleID)

```

##Giant relative abundance table for Meagan
```{r All relative abundance table}
g_taxa_df <- classify_df %>% 
  dplyr::filter(., Rank == "species") %>%
  droplevels.data.frame() %$%
  getTaxonomy(unique(TaxID),taxaNodes,taxaNames) %>%
  as.data.frame() %>%
  dplyr::mutate(., TaxID = rownames(.)) %$%
  dplyr::mutate(., TaxID = as.factor(trimws(TaxID, which = "both")))

classify_genus_df <- merge(classify_df, g_taxa_df, "TaxID")

sm_gen_df <- dplyr::select(classify_genus_df, Sample, CladeNumberReads, TaxonNumberReads, species, Name) 

  
abundance_df <-sm_gen_df %>% 
  dplyr::mutate(TotalReads = (CladeNumberReads + 
                                TaxonNumberReads)) %>% 
  dplyr::select(c("TotalReads", "Name", "Sample")) %$%
  dplyr::mutate(., Name = as.character(Name)) %$%
  reshape2::acast(., Name ~ Sample, 
                  value.var = "TotalReads", 
                  fun.aggregate = sum) 


#Check that things will add up to 1
colSums(prop.table(abundance_df,2))

all_abundance <- prop.table(abundance_df,2)
#Check again
colSums(all_abundance,2)


all_abundance1 <- as.data.frame(all_abundance[,!colnames(all_abundance) %in% IDdrops])

all_abundance <- as.data.frame(all_abundance1[,!colnames(all_abundance1) %in% newIDdrops])

#all_abundance <- as.data.frame(all_abundance2[,!colnames(all_abundance2) %in% newIDdrops2])

all_abundance_t <- as.data.frame(t(all_abundance))
#all_abundance_t <- as.data.frame(t(all_abundance1))


```

```{r Data Preprocessing}
#Remove rare features

all_abundance_t_nonzero_counts <- apply(all_abundance_t, 1, function(y) sum(length(which(y > 0))))

hist(all_abundance_t_nonzero_counts, breaks=100, col="grey", main="", ylab="Number of Taxa", xlab="Number of Non-Zero Values")

#Based on the histogram, remove non-zero values in some percentage of the samples (e.g., .20 of samples) 
#abundance_rare_removed <- remove_rare(table=all_abundance_t, cutoff_pro=0.50)
all_abundance_t <- t(all_abundance_t)
abundance_rare_removed <- remove_rare(table=all_abundance_t, cutoff_pro = 0.5)
#head(abundance_rare_removed)
#dim(abundance_rare_removed)
#dim(all_abundance_t)
#rownames(abundance_rare_removed)
#renormalize table so that each sample's column sums to 100 
abundance_norm <- sweep(abundance_rare_removed, 2, colSums(abundance_rare_removed) , '/')*100

#head(abundance_norm)
#dim(abundance_norm)
#rownames(abundance_norm)

#Transform data into normalized z-scores by subtracting each sample's mean (center) and then dividing by the sample's standard deviation (scale).
abundance_scaled <- scale(abundance_norm, center = TRUE, scale = TRUE)
#Prep input tables for classification of state
abundance_scaled <- as.data.frame(t(abundance_scaled)) #double check that rows are sample id and taxa are columns
#abundance_scaled <- t(abundance_scaled)

abundance_scaled$BothCommensalPathogen <- metadata_IDs_filtered[rownames(abundance_scaled), 'BothCommensalPathogen']
abundance_scaled$BothCommensalPathogen <- as.integer(abundance_scaled$BothCommensalPathogen)
#abundance_scaled <- as.data.frame(abundance_scaled)
#abundance_scaled$IL.5[6] <- "No"
#abundance_scaled$IL.5[12] <- "No"
#abundance_scaled$IL.5[54] <- "No"
#Add positive control
abundance_scaled_poscntrl <- abundance_scaled %>%
  mutate(fake_taxon = add_noise(BothCommensalPathogen))

#Drop unclassified
abundance_scaled_poscntrl$unclassified <- NULL
dim(abundance_scaled_poscntrl)
str(abundance_scaled_poscntrl)
abundance_scaled_poscntrl$BothCommensalPathogen <- as.factor(abundance_scaled_poscntrl$BothCommensalPathogen)
str(abundance_scaled_poscntrl$BothCommensalPathogen)

```
Running RFC
```{r}
#Guidelines: use half caases, match with controls. 
#table(abundance_scaled_poscntrl$BothCommensalPathogen)
#1/2 pos = 13. Each group =13
#Set the random seed so that models will be reproducible.
set.seed(151)  

#parallelize  job
cl <- makePSOCKcluster(20)
registerDoParallel(cl)

colnames(abundance_scaled_poscntrl) <- make.names(colnames(abundance_scaled_poscntrl))

#abundance_scaled_poscntrl$Lactose_Binary <- gsub(1, "LNP", abundance_scaled_poscntrl$Lactose_Binary)
#abundance_scaled_poscntrl$Lactose_Binary <- gsub(2, "LP", abundance_scaled_poscntrl$Lactose_Binary)
#abundance_scaled_poscntrl$Lactose_Binary <- as.factor(abundance_scaled_str(poscntrl$Lactose_Binary)
#abundance_scaled_poscntrl <- as.matrix()

RF_abundance_scaled_poscntrl_DS = randomForest(BothCommensalPathogen ~ ., abundance_scaled_poscntrl,ntree=5001,sampsize=c(13,13),strata=abundance_scaled_poscntrl$BothCommensalPathogen)

#RF_asv_scaled_regression = randomForest(Lactose_Binary ~ ., abundance_scaled_poscntrl,ntree=5001,importance=TRUE, proximities=TRUE) #poscontrl included 
#poscontrl, downsample

print(RF_abundance_scaled_poscntrl_DS)

par(mfrow=c(1,2))
RF_state_classify_imp <- as.data.frame(RF_abundance_scaled_poscntrl_DS$importance )
RF_state_classify_imp$features <- rownames( RF_state_classify_imp )
RF_state_classify_imp_sorted <- arrange( RF_state_classify_imp  , desc(MeanDecreaseGini)  )
RF_state_classify_imp_sorted <- RF_state_classify_imp_sorted[!grepl("fake_taxon", RF_state_classify_imp_sorted$features),] #Drop fake taxa fake_taxon


barplot(RF_state_classify_imp_sorted$MeanDecreaseGini, ylab="Mean Decrease in Accuracy (Variable Importance)", main="RF Classification Variable Importance Distribution")

barplot(RF_state_classify_imp_sorted[1:13,"MeanDecreaseGini"], names.arg=RF_state_classify_imp_sorted[1:13,"features"] , ylab="Mean Decrease in Accuracy (Variable Importance)", las=2, ylim=c(0,2), main="Classification RF with Positive Control")  

Parasite_class <- data.frame(Feature = RF_state_classify_imp_sorted[1:13,"features"], Gini = RF_state_classify_imp_sorted[1:13,"MeanDecreaseGini"])

p <- ggplot(Parasite_class, aes(x= reorder(Feature, Gini), y = Gini)) +
ggtitle("Entamoeba Commensal & Pathogenic RF Classification") + 
ylab("Mean Decrease in Gini (Variable Importance)") + 
xlab("Feature") +
geom_bar(stat="identity") + theme_bw()

p + coord_flip()

stopCluster(cl)
```

```{r}
BA_Porphy <- ggplot(data = abundance_scaled_poscntrl, aes(x = BothCommensalPathogen, y = X..........Blastocystis.hominis)) + 
  geom_boxplot()+
  geom_point(position = position_jitter(width = 0.25), na.rm = TRUE) +
  #facet_grid(. ~ Nematode) +
  geom_point(aes(color = factor(BothCommensalPathogen))) +
 scale_x_continuous(breaks = c(1, 15, 30, 45, 60, 75)) +
  #ylim(0,4.5) +
  geom_smooth(method = lm) +
  geom_point(color='#453781FF')  +
  labs(title = 'Normalized Abundances of RFC taxa and IL5',y= 'Porphyromonas asaccharolytica', x =   'IL.5') +
  theme_bw()

BA_Porphy <- ggplot(data = abundance_scaled_poscntrl, aes(x = IL.5, y = X....................Porphyromonas.asaccharolytica)) + 
  #geom_boxplot()+
  geom_point(position = position_jitter(width = 0.25), na.rm = TRUE) +
  #facet_grid(. ~ Nematode) +
  geom_point(aes(color = factor(IL.5))) +
 scale_x_continuous(breaks = c(1, 15, 30, 45, 60, 75)) +
  #ylim(0,4.5) +
  geom_smooth(method = lm) +
  geom_point(color='#453781FF')  +
  labs(title = 'Normalized Abundances of RFC taxa and IL5',y= 'Porphyromonas asaccharolytica', x =   'IL.5') +
  theme_bw()

theme_aa_scatter <- theme_bw(base_size = 12) +
  theme(panel.grid.minor = element_blank(),
        strip.background = element_rect(fill = "white", color = "black"),
        axis.text = element_text(size = rel(1.1), color = "black"),
        axis.title = element_text(size = rel(1.1)))

newtest <- abundance_scaled_poscntrl %>% select(
`X..........Blastocystis.hominis`,
`X....................Flavobacterium.magnum`,
`X....................Erysipelotrichaceae.bacterium.SG0102`,
`X..................Trueperella.pyogenes`,
`X..................Staphylococcus.aureus`,
`X................Shigella.dysenteriae`,
`X................Rhodobacter.sphaeroides`,
`X..........bacterium`,
`X..................Anoxybacillus.kamchatkensis`,
`X..................Enterobacter.cloacae`,
BothCommensalPathogen)


new <- melt(newtest, "BothCommensalPathogen")

theme_scatter <- theme_bw(base_size = 24) + 
  theme(
  panel.grid.major.x = element_blank(),
  panel.grid.major.y = element_blank(),
  panel.grid.minor.x = element_blank(),
  panel.grid.minor.y = element_blank(),
  line = element_line(size=1),
  text = element_text(color = "black"),
  axis.text.x = element_text(color="black"),
  axis.text.y = element_text(color="black"),
  strip.background=element_rect(fill="white", color="white")
)

mt <- ggplot(new, aes(variable, value, fill = factor(BothCommensalPathogen))) +
  geom_boxplot() + 
  theme_classic(base_size = 20) +
  theme(axis.text.x=element_text(angle=45, hjust=1), strip.background=element_rect(fill="white", color="white"))  + theme_scatter + labs(fill = "Entamoeba Positive", x = "Shotgun Sequencing taxon", y = "Relative  Abundance (z-score)")




a <- pairwise.wilcox.test(abundance_scaled_poscntrl$X..........Blastocystis.hominis, abundance_scaled_poscntrl$BothCommensalPathogen)
b<- pairwise.wilcox.test(abundance_scaled_poscntrl$X....................Flavobacterium.magnum, abundance_scaled_poscntrl$BothCommensalPathogen)
c<- pairwise.wilcox.test(abundance_scaled_poscntrl$X....................Erysipelotrichaceae.bacterium.SG0102, abundance_scaled_poscntrl$BothCommensalPathoge)
d<- pairwise.wilcox.test(abundance_scaled_poscntrl$X..................Trueperella.pyogenes, abundance_scaled_poscntrl$BothCommensalPathogen)
e<- pairwise.wilcox.test(abundance_scaled_poscntrl$X..................Staphylococcus.aureus, abundance_scaled_poscntrl$BothCommensalPathogen)
f<- pairwise.wilcox.test(abundance_scaled_poscntrl$X................Shigella.dysenteriae, abundance_scaled_poscntrl$BothCommensalPathogen)
g<- pairwise.wilcox.test(abundance_scaled_poscntrl$X................Rhodobacter.sphaeroides, abundance_scaled_poscntrl$BothCommensalPathogen)
h<- pairwise.wilcox.test(abundance_scaled_poscntrl$X..........bacterium, abundance_scaled_poscntrl$BothCommensalPathogen)
i<- pairwise.wilcox.test(abundance_scaled_poscntrl$X..................Anoxybacillus.kamchatkensis, abundance_scaled_poscntrl$BothCommensalPathogen)
j<- pairwise.wilcox.test(abundance_scaled_poscntrl$X..................Enterobacter.cloacae, abundance_scaled_poscntrl$BothCommensalPathogen)

third <- data.frame(taxon= c(a$data.name, b$data.name, c$data.name, d$data.name, e$data.name, f$data.name, g$data.name, h$data.name, i$data.name, j$data.name),
                    pval=c(a$p.value,b$p.value,c$p.value,d$p.value,e$p.value,f$p.value,g$p.value,h$p.value,i$p.value,j$p.value))
third$fdr <- NA

third$fdr <-p.adjust(third$pval, method="fdr")



lm(formula = abundance_scaled_poscntrl$BothCommensalPathogen ~ abundance_scaled_poscntrl$X..........Blastocystis.hominis)
#porphyro goes up in IL5

Prev <- grep('Prevotella.stercorea', RF_state_classify_imp_sorted$features, value = TRUE) %>% as.data.frame()
colnames(Prev) <- "features"
Prev_gini <- left_join(Prev, RF_state_classify_imp_sorted, by = 'features')
Prev_gini <- mutate(Prev_gini, TotalGini = sum(`%IncMSE`)) %>% View()

Rum <- grep('Ruminococcus.bromii', RF_state_classify_imp_sorted$features, value = TRUE) %>% as.data.frame()
colnames(Rum) <- "features"
Rum_gini <- left_join(Rum, RF_state_classify_imp_sorted, by = 'features')
Rum_gini <- mutate(Rum_gini, TotalGini = sum(`%IncMSE`)) %>% View()

Suc <- grep('Succinivibrio', RF_state_classify_imp_sorted$features, value = TRUE) %>% as.data.frame()
colnames(Suc) <- "features"
Suc_gini <- left_join(Suc, RF_state_classify_imp_sorted, by = 'features')
Suc_gini <- mutate(Suc_gini, TotalGini = sum(`%IncMSE`)) %>% View()
```


Build RFC on just pathogenic Entamoeba
```{r}
abundance_scaled$PathogenOnly <- metadata_IDs_filtered[rownames(abundance_scaled), 'PathogenOnly']
abundance_scaled$PathogenOnly <- as.integer(abundance_scaled$PathogenOnly)
#abundance_scaled <- as.data.frame(abundance_scaled)
#abundance_scaled$IL.5[6] <- "No"
#abundance_scaled$IL.5[12] <- "No"
#abundance_scaled$IL.5[54] <- "No"
#Add positive control
abundance_scaled_poscntrl <- abundance_scaled %>%
  mutate(fake_taxon = add_noise(PathogenOnly))

#Drop unclassified
abundance_scaled_poscntrl$unclassified <- NULL
abundance_scaled_poscntrl$BothCommensalPathogen <- NULL

dim(abundance_scaled_poscntrl)
str(abundance_scaled_poscntrl)
abundance_scaled_poscntrl$PathogenOnly <- as.factor(abundance_scaled_poscntrl$PathogenOnly)
str(abundance_scaled_poscntrl$PathogenOnly)


```
Running RFC
```{r}
#Guidelines: use half caases, match with controls. 
#table(abundance_scaled_poscntrl$BothCommensalPathogen)
#1/2 pos = 3. Each group =3
#Set the random seed so that models will be reproducible.
set.seed(151)  

#parallelize  job
cl <- makePSOCKcluster(20)
registerDoParallel(cl)

colnames(abundance_scaled_poscntrl) <- make.names(colnames(abundance_scaled_poscntrl))

#abundance_scaled_poscntrl$Lactose_Binary <- gsub(1, "LNP", abundance_scaled_poscntrl$Lactose_Binary)
#abundance_scaled_poscntrl$Lactose_Binary <- gsub(2, "LP", abundance_scaled_poscntrl$Lactose_Binary)
#abundance_scaled_poscntrl$Lactose_Binary <- as.factor(abundance_scaled_str(poscntrl$Lactose_Binary)
#abundance_scaled_poscntrl <- as.matrix()

RF_abundance_scaled_poscntrl_DS = randomForest(PathogenOnly ~ ., abundance_scaled_poscntrl,ntree=5001,sampsize=c(3,3),strata=abundance_scaled_poscntrl$PathogenOnly)

#RF_asv_scaled_regression = randomForest(Lactose_Binary ~ ., abundance_scaled_poscntrl,ntree=5001,importance=TRUE, proximities=TRUE) #poscontrl included 
#poscontrl, downsample

print(RF_abundance_scaled_poscntrl_DS)

par(mfrow=c(1,2))
RF_state_classify_imp <- as.data.frame(RF_abundance_scaled_poscntrl_DS$importance )
RF_state_classify_imp$features <- rownames( RF_state_classify_imp )
RF_state_classify_imp_sorted <- arrange( RF_state_classify_imp  , desc(MeanDecreaseGini)  )
RF_state_classify_imp_sorted <- RF_state_classify_imp_sorted[!grepl("fake_taxon", RF_state_classify_imp_sorted$features),] #Drop fake taxa fake_taxon


barplot(RF_state_classify_imp_sorted$MeanDecreaseGini, ylab="Mean Decrease in Accuracy (Variable Importance)", main="RF Classification Variable Importance Distribution")

barplot(RF_state_classify_imp_sorted[1:10,"MeanDecreaseGini"], names.arg=RF_state_classify_imp_sorted[1:10,"features"] , ylab="Mean Decrease in Accuracy (Variable Importance)", las=2, ylim=c(0,2), main="Classification RF with Positive Control")  

Parasite_class <- data.frame(Feature = RF_state_classify_imp_sorted[1:10,"features"], Gini = RF_state_classify_imp_sorted[1:10,"MeanDecreaseGini"])

p <- ggplot(Parasite_class, aes(x= reorder(Feature, Gini), y = Gini)) +
ggtitle("Entamoeba Pathogenic RF Classification") + 
ylab("Mean Decrease in Gini (Variable Importance)") + 
xlab("Feature") +
geom_bar(stat="identity") + theme_bw()

p + coord_flip()

stopCluster(cl)
```
Build RFC on just commensal Entamoeba
```{r}
abundance_scaled$CommensalOnly <- metadata_IDs_filtered[rownames(abundance_scaled), 'CommensalOnly']
abundance_scaled$CommensalOnly <- as.integer(abundance_scaled$CommensalOnly)
#abundance_scaled <- as.data.frame(abundance_scaled)
#abundance_scaled$IL.5[6] <- "No"
#abundance_scaled$IL.5[12] <- "No"
#abundance_scaled$IL.5[54] <- "No"
#Add positive control
abundance_scaled_poscntrl <- abundance_scaled %>%
  mutate(fake_taxon = add_noise(CommensalOnly))

#Drop unclassified
abundance_scaled_poscntrl$unclassified <- NULL
abundance_scaled_poscntrl$PathogenOnly <- NULL

dim(abundance_scaled_poscntrl)
str(abundance_scaled_poscntrl)
abundance_scaled_poscntrl$CommensalOnly <- as.factor(abundance_scaled_poscntrl$CommensalOnly)
str(abundance_scaled_poscntrl$CommensalOnly)
table(abundance_scaled_poscntrl$CommensalOnly)

```
Running RFC
```{r}
#Guidelines: use half caases, match with controls. 
#table(abundance_scaled_poscntrl$BothCommensalPathogen)
#1/2 pos = 8. Each group =8
#Set the random seed so that models will be reproducible.
set.seed(151)  

#parallelize  job
cl <- makePSOCKcluster(20)
registerDoParallel(cl)

colnames(abundance_scaled_poscntrl) <- make.names(colnames(abundance_scaled_poscntrl))

#abundance_scaled_poscntrl$Lactose_Binary <- gsub(1, "LNP", abundance_scaled_poscntrl$Lactose_Binary)
#abundance_scaled_poscntrl$Lactose_Binary <- gsub(2, "LP", abundance_scaled_poscntrl$Lactose_Binary)
#abundance_scaled_poscntrl$Lactose_Binary <- as.factor(abundance_scaled_str(poscntrl$Lactose_Binary)
#abundance_scaled_poscntrl <- as.matrix()

RF_abundance_scaled_poscntrl_DS = randomForest(CommensalOnly ~ ., abundance_scaled_poscntrl,ntree=5001,sampsize=c(8,8),strata=abundance_scaled_poscntrl$CommensalOnly)

#RF_asv_scaled_regression = randomForest(Lactose_Binary ~ ., abundance_scaled_poscntrl,ntree=5001,importance=TRUE, proximities=TRUE) #poscontrl included 
#poscontrl, downsample

print(RF_abundance_scaled_poscntrl_DS)

par(mfrow=c(1,2))
RF_state_classify_imp <- as.data.frame(RF_abundance_scaled_poscntrl_DS$importance )
RF_state_classify_imp$features <- rownames( RF_state_classify_imp )
RF_state_classify_imp_sorted <- arrange( RF_state_classify_imp  , desc(MeanDecreaseGini)  )
RF_state_classify_imp_sorted <- RF_state_classify_imp_sorted[!grepl("fake_taxon", RF_state_classify_imp_sorted$features),] #Drop fake taxa fake_taxon


barplot(RF_state_classify_imp_sorted$MeanDecreaseGini, ylab="Mean Decrease in Accuracy (Variable Importance)", main="RF Classification Variable Importance Distribution")

barplot(RF_state_classify_imp_sorted[1:10,"MeanDecreaseGini"], names.arg=RF_state_classify_imp_sorted[1:10,"features"] , ylab="Mean Decrease in Accuracy (Variable Importance)", las=2, ylim=c(0,2), main="Classification RF with Positive Control")  

Parasite_class <- data.frame(Feature = RF_state_classify_imp_sorted[1:10,"features"], Gini = RF_state_classify_imp_sorted[1:10,"MeanDecreaseGini"])

p <- ggplot(Parasite_class, aes(x= reorder(Feature, Gini), y = Gini)) +
ggtitle("Entamoeba Commensal RF Classification") + 
ylab("Mean Decrease in Gini (Variable Importance)") + 
xlab("Feature") +
geom_bar(stat="identity") + theme_bw()

p + coord_flip()

stopCluster(cl)
```

```{r}
shotgun_metadata_cameroon <- read_delim("/media/THING2/louis/15_RubelCameroonShotgun/consign-kegg-r-ator/shotgun_metadata_cameroon_5119.csv", ",", escape_double = FALSE, trim_ws = TRUE)

#shotgun_metadata_cameroon[is.na(shotgun_metadata_cameroon)] <- "Unknown" 
CM_metadata <- shotgun_metadata_cameroon[!(shotgun_metadata_cameroon$ants_binary== "Unknown"),]

                                              
BA_Porphy <- ggplot(data = abundance_scaled_poscntrl, aes(x = IL.5, y = X....................Porphyromonas.asaccharolytica)) + 
  #geom_boxplot()+
  geom_point(position = position_jitter(width = 0.25), na.rm = TRUE) +
  #facet_grid(. ~ Nematode) +
  geom_point(aes(color = factor(IL.5))) +
 scale_x_continuous(breaks = c(1, 15, 30, 45, 60, 75)) +
  #ylim(0,4.5) +
  geom_smooth(method = lm) +
  geom_point(color='#453781FF')  +
  labs(title = 'Normalized Abundances of RFC taxa and IL5',y= 'Porphyromonas asaccharolytica', x =   'IL.5') +
  theme_bw()

CM_metadata$IL.13 <- as.numeric(CM_metadata$IL.13)

x <- ggplot(CM_metadata, aes(x = factor(ants_binary), y = IL.13)) +
geom_boxplot(na.rm=TRUE) + 
  theme_bw()
