
```{r DeNOVO OTU code, echo=FALSE}
# Assign sample mapping file
s <- read_qiime_mapping_file("P55_mapping_Rbx_UPenn.txt")

# Replicate sample
s_original <- subset(s, Sample.Pooling %in% "Original")

# Assign OTU table
o <- read_qiime_otu_table("otu/classic_otu_table.txt")

# Metadata in the form of truncated green genes assignments
md <- sub("(; [kpcofgs]__)+$", "", o$metadata, perl=T)

# Assignments data-frame
adf <- split_assignments(md)
a <- simplify_assignments(adf)

# Arrange s by StudyGroup, then by SubjectID, then by Age_bin
s <- arrange(s, StudyGroup, SubjectID, Age_bin)

# Check for invalid sample IDs
problem_ids <- setdiff(s$SampleID, colnames(o$counts))
if (length(problem_ids ) > 0) stop (simpleError(paste("id mismatch found", problem_ids, collapse=" ")))

# Assignment counts per SampleID
cts <- o$counts[,s$SampleID]
```


## --------------------------------------------------------

```{r echo=FALSE, message=FALSE}
unw <- read_qiime_distmat("beta_diversity/unweighted_unifrac_dm.txt")
wtu <- read_qiime_distmat("beta_diversity/weighted_normalized_unifrac_dm.txt")
```
## Beta diversity analyses

```{r}
mapping_subset <- s_rectal_S004
metadata_col <- "BodySite"
mapping_subset$Period <- factor(mapping_subset$Period)
```

### Presence/Absence Unifrac plots, colored by `r metadata_col`
```{r}
unw_pcoa <- pcoa_1way(mapping_subset, unw, var=metadata_col)
unw_pcoa$plot # + scale_color_manual(values=c("JK"="blue","AM"="green","BM"="red"))
```
##### Presence/Absence test for effect (group centroids), by `r metadata_col`
```{r}
udf <- droplevels(unw_pcoa$df)
adonis(unw_pcoa$distances ~ BodySite, data=udf)
```

### Abundance weighted Unifrac plots, colored by `r metadata_col`

```{r}
wtu_pcoa <- pcoa_1way(mapping_subset, wtu, var=metadata_col)
wtu_pcoa$plot #+ scale_color_manual(values=c("JK"="blue","AM"="green","BM"="red"))
```
##### Weighted test for effect (group centroids), by `r metadata_col`
```{r}
wdf <- droplevels(wtu_pcoa$df)
adonis(wtu_pcoa$distances ~ BodySite, data=wdf)
```

## Differing taxa by Kruskall-Wallis between `r metadata_col`
No Significant Differences (underpowered)
```{r DifferingTaxonAbundance_figure, fig.height=9, echo=FALSE}
#test_taxon_abundance(mapping_subset, cts, metadata_col, a=a )
``` 
