---
title: "Rebiotix study part 3"
author: "Aubrey Bailey"
date: "01/28/2015"
output: html_document
---
##16S analysis of longitudinal incubated artifical fecal transplant donations

```{r setup, echo=FALSE}
#Creates figure image files in Knit folder 
library(knitr)
opts_chunk$set(
  tidy=FALSE,
  cache=F,
  echo=F,
  dpi=100,
  fig.width=6,
  fig.path="Knit/figure/",
  cache.path="Knit/cache/",
  dev=c("png", "pdf", "postscript"))
```

```{r load scripts, echo=FALSE, message=FALSE}
#Load scripts for 16S analysis
library(qiimer)
library(ape)
library(vegan)
library(ggplot2)
library(plyr)
library(reshape2)
library(kylemisc)
```

# make fake OTU table
```{r}
num_groups <- 3
subj_per_group <- 6
num_tests <- 2
cts  <- simStudy(group_size_vector = c(subj_per_group, num_groups, num_tests),
                otu_number=10,
                sequence_depth=1000,
                rare_depth = 0.8,
#                 effect_range = seq(0,0.3,length.out=100)
)
head(cts,1)


#unifracs <- GUniFrac(otu.tab.rff, throat.tree, alpha=c(0, 0.5, 1))$unifracs
#dw <- unifracs[, , "d_1"]
```

```{r DeNOVO OTU code, echo=FALSE}
# Assign sample mapping file
s <- read_qiime_mapping_file("P55_mapping_Rbx_UPenn.txt")

# Replicate sample
s_original <- subset(s, Sample.Pooling %in% "Original")

# Assign OTU table
o <- read_qiime_otu_table("otu/classic_otu_table.txt")

# Metadata in the form of truncated green genes assignments
md <- sub("(; [kpcofgs]__)+$", "", o$metadata, perl=T)

# Assignments data-frame
adf <- split_assignments(md)
a <- simplify_assignments(adf)

# Arrange s by StudyGroup, then by SubjectID, then by Age_bin
s <- arrange(s, StudyGroup, SubjectID, Age_bin)

# Check for invalid sample IDs
problem_ids <- setdiff(s$SampleID, colnames(o$counts))
if (length(problem_ids ) > 0) stop (simpleError(paste("id mismatch found", problem_ids, collapse=" ")))

# Assignment counts per SampleID
cts <- o$counts[,s$SampleID]
```

## --------------------------------------------------------
### Sample structure for original stool design  
```{r}
table(s_original$Collection_Day, s_original$Replicate_Number, dnn=c("Day","Replicate"))
```
### Sample Structure for pooled stool deisgn
```{r}
table(s$Collection_Day, s$Drug_SubstanceSample_ID, dnn=c("Day","Batch"))
```
## --------------------------------------------------------

```{r}
read_counts <- read.csv("library/split_library_log.txt", sep="\t", skip=15, strip.white = FALSE, blank.lines.skip = TRUE,  header=FALSE, nrows = length(file) -1, col.names=c("SampleID", "Read Counts"))

read_counts <- read_counts[order(read_counts$SampleID), ]
read_counts[1:(nrow(read_counts)-2),]
```

##---------------------------------------------------------

## Alpha Diversity

### Alpha diversity at enrollment (t1)

Prior to study enrollment, infants in the Formula A and B groups were on a variety of formula diets.

```{r alpha_diversity,echo=FALSE}
ad <- read_qiime_rarefaction("alpha_diversity/alpha_div_collated/observed_species.txt")
```

```{r alpha_diversity_join, echo=FALSE}
rs <- rarefaction_stats(ad)
rs_t1_joined <- merge(s_original, rs, by="SampleID")
rs_joined <- merge(s, rs, by="SampleID")
```

```{r fig.width=6, fig.height=4, echo=FALSE}
ggplot(rs_t1_joined, aes(x=sequences_per_sample, y=diversity.mean, ymin=0, ymax=0, group=SampleID, color=Collection_Day)) + 
  stat_smooth(method = "loess", formula = y ~ x,  se = TRUE, alpha=.3) +
  xlab("Sequences Per Sample") +
  ylab("Mean Observed Species") +
  ggtitle("Alpha Diversity Rarefaction Curves\n Observed Species")
  #geom_errorbar(aes(ymin=diversity.mean - diversity.sd, ymax=diversity.mean + diversity.sd)) +
  #geom_point( shape=21, size=3) + # 21 is filled circle
  #scale_color_manual(values=c("BREASTFED"="red","FORMULA A"  ="blue","FORMULA B"="green"))
```

```{r, fig.width=6, fig.height=4, echo=FALSE}
ggplot(rs_t1_joined, aes(x=sequences_per_sample, y=diversity.mean, ymin=0, ymax=0, group=SampleID, color=StudyGroup)) + 
  stat_smooth(method = "loess", formula = y ~ x,  se = TRUE, alpha=.3) +
  xlab("Sequences Per Sample") +
  ylab("Mean Observed Species") +
  ggtitle("Alpha Diversity Rarefaction Curves\n Observed Species") +

```


## --------------------------------------------------------

### De Novo OTU Heatmap

```{r overall_heatmap_annotations, echo=FALSE}
annotations <- s[,c("SubjectID","StudyGroup","Age_bin")]
rownames(annotations) <- s$SampleID
```

```{r simOTU_Heatmap, fig.width=15, fig.height=10, echo=FALSE}
otu_heatmap(
  cts, rownames(cts), threshold=200, 
  color = saturated_rainbow(max(colSums(cts)), saturation_limit = 0.4), 
  cluster_cols = FALSE, cluster_rows = FALSE, fontsize_col = 12, fontsize_row = 12, 
  legend = TRUE, cellheight=12, cellwidth=12)
```

```{r DeNovoOTU_Heatmap, fig.width=15, fig.height=10, echo=FALSE}
# otu_heatmap(
#   cts, a, annotation=annotations, threshold=1000, 
#   color = saturated_rainbow(max(colSums(cts)), saturation_limit = 0.4), 
#   cluster_cols = FALSE, cluster_rows = FALSE, fontsize_col = 12, fontsize_row = 12, 
#   legend = TRUE, cellheight=12, cellwidth=12)
```

```{r subject_heatmap_annotations, echo=FALSE}
annotations <- s[,c("SubjectID","InfantDiet","Age_bin")]
rownames(annotations) <- s$SampleID
```

### Heatmaps by Subject
3-4 samples were collected from each of the 15 subjects at separate timepoints, over a span of approximately 100 days.

```{r Heatmap_by_SubjectID, fig.width=8, fig.height=7, echo=FALSE}
ann_colors <- list(
  InfantDiet=c("\"PHF, CMF\""="purple","BREASTFED"="red","CMF"="orange","FORMULA A"="blue","FORMULA B"="green","PHF"="yellow","SOY"="gray"), 
  Age_bin=c("1"="red","2"="orange","3"="green","4"="blue"))                                                                                     

ann_colors$Age_bin <- as.factor(ann_colors$Age_bin)
s$Age_bin <- as.factor(s$Age_bin)

#Heatmap maker with loop
d_ply(s, .(SubjectID), function (x) {
   title <- paste(x[1,"SubjectID"])
   otu_heatmap(
     cts[,x$SampleID], a, annotation=annotations[,c("Age_bin","InfantDiet")], threshold=400, main=title, annotation_colors=ann_colors, 
     color = saturated_rainbow(max(colSums(cts)), saturation_limit = 0.40), 
    cluster_cols = FALSE, cluster_rows = FALSE, fontsize_col = 10, fontsize_row = 10, 
    legend = TRUE, cellheight=12, cellwidth=10)
})
```

### Heatmaps by Age Bin (1-4)

```{r Heatmap_by_Age_bin, fig.width=10, fig.height=8, echo=FALSE}
# Arrange s by InfantDiet
# s <- arrange(s, InfantDiet, SubjectID)

#Heatmap maker with loop

ann_colors <- list(
  InfantDiet=c("BREASTFED"="red","FORMULA A"="blue","FORMULA B"="green","PHF"="yellow","CMF"="orange","\"PHF, CMF\""="purple","SOY"="gray")
)                                                                                         

d_ply(s, .(Age_bin), function (x) {
   title <- paste(x[1,"Age_bin"])
   otu_heatmap(
     cts[,x$SampleID], a, annotation=annotations[
       ,c("SubjectID","InfantDiet")], threshold=400, main=title, annotation_colors=ann_colors, 
     color = saturated_rainbow(max(colSums(cts)), saturation_limit = 0.40), 
    cluster_cols = FALSE, cluster_rows = FALSE, fontsize_col = 10, fontsize_row = 10, 
    legend = TRUE, cellheight=12, cellwidth=10)
})
```

## --------------------------------------------------------

```{r echo=FALSE, message=FALSE}
#unw <- read_qiime_distmat("beta_diversity/unweighted_unifrac_dm.txt")
#wtu <- read_qiime_distmat("beta_diversity/weighted_unifrac_dm.txt")
wtu <- calcWJstudy(otu_table = cts)
groupNames(wtu)
```

```{r pcoa_1way, echo=FALSE, message=FALSE}
#' Perform a PCoA analysis of one variable
#' @param s The sample table
#' @param d A distance matrix.  The matrix will be trimmed to include only the samples in `s`.
#' @param var The variable of interest (string)
#' @param strata If data is stratified, the name of the column containing the strata
#' @param ci Confidence interval to plot ellipse around. NULL if not plotted.
pcoa_1way <- function(s, d, var="StudyGroup", strata=NULL, ci=NULL) {
  # Select columns of interest from distance matrix
  distances <- dist_subset(d, s$SampleID)

  res <- list()
  
  # Run the actual PCoA ordination
  res$pcoa <- pcoa(distances)
  
  # Save distances and sample info back into result
  res$distances <- distances
  res$df <- cbind(s, res$pcoa$vectors[,1:2])
  
  # Make a simple plot, save in result
  res$plot <- ggplot(res$df, aes_string(x="Axis.1", y="Axis.2", color=var)) + 
    geom_point() + theme_classic()
  # scale_colour_brewer(palette="Set1") +
  # geom_text(aes(label = SampleID )) +  ( add , label=s$SampleID to aes_string if using this)

    
  if (!is.null(ci)) {
    res$plot <- res$plot + stat_ellipse(level=ci, linetype=2)
  }
  
  # Run an adonis test, save in result
  adonis_formula <- formula(paste("distances ~", var))
  strata_arg <- if (is.null(strata)) NULL else res$df[[strata]]
  res$test <- adonis(adonis_formula, data=s, strata=strata_arg)
  
  # Fix the test results:
  # Print the actual formula instead of "adonis_formula"
  # Replace the strata, too, for readability
  adonis_call <- as.list(res$test$call)
  adonis_call$formula <- adonis_formula
  # Have to do some R metaprogramming to make this look right...
  adonis_call$strata <- if (is.null(strata)) NULL else substitute(s$strata, list(strata=strata))
  res$test$call <- as.call(adonis_call)

  class(res) <- "pcoa_1way"
  res
}

print.pcoa_1way <- function (x) {
  print(x$plot)
#  print(x$test)
}
```

```{r pcoa_1way_continuous, echo=FALSE}
#' Perform a PCoA analysis of one variable
#' @param s The sample table
#' @param d A distance matrix.  The matrix will be trimmed to include only the samples in `s`.
#' @param var The variable of interest (string)
#' @param strata If data is stratified, the name of the column containing the strata
#' @param ci Confience interval to plot elipse around. NULL if not ploted.
pcoa_1way_continuous <- function(s, d, var="study_group", strata=NULL, ci=NULL) {
  # Select columns of interest from distance matrix
  distances <- dist_subset(d, s$SampleID)

  res <- list()
  
  # Run the actual PCoA ordination
  res$pcoa <- pcoa(distances)
  
  # Save distances and sample info back into result
  res$distances <- distances
  res$df <- cbind(s, res$pcoa$vectors[,1:3])
  
  # Make a simple plot, save in result
  res$plot <- ggplot(res$df, aes_string(x="Axis.1", y="Axis.2",color=var)) + geom_point() + theme_classic() + scale_colour_gradientn(colours = rainbow(2))
  
  if (!is.null(ci)) {
    res$plot <- res$plot + stat_ellipse(level=ci, linetype=2)
  }
  
  # Run an adonis test, save in result
  adonis_formula <- formula(paste("distances ~", var))
  strata_arg <- if (is.null(strata)) NULL else res$df[[strata]]
  res$test <- adonis(adonis_formula, data=s, strata=strata_arg)
  
  # Fix the test results:
  # Print the actual formula instead of "adonis_formula"
  # Replace the strata, too, for readability
  adonis_call <- as.list(res$test$call)
  adonis_call$formula <- adonis_formula
  # Have to do some R metaprogramming to make this look right...
  adonis_call$strata <- if (is.null(strata)) NULL else substitute(s$strata, list(strata=strata))
  res$test$call <- as.call(adonis_call)

  class(res) <- "pcoa_1way_continuous"
  res
}

print.pcoa_1way_continuous <- function (x) {
  print(x$plot)
#  print(x$test)
}
```

```{r test_taxon_abundance, echo=FALSE}
#' Test taxa for differences in abundance
#' @param s A data frame of sample info.
#' @param cts A matrix of OTU counts (rows) for each sample (columns).
#' @param g A factor of sample groups to be tested.
#' @param a A chearacter vector of OTU assignments.
test_taxon_abundance <- function (s, cts, var="study_group", a=NULL, min_fraction=0.5) {
  # Limit OTU table to samples in provided data frame
  cts <- cts[, s$SampleID]
  cts <- rowsum(cts, as.character(a))
  
  # Matrix of OTU proportions
  props <- sweep(cts, 2, colSums(cts), `/`)

  # Empty result
  res <- list()

  # Detect OTUs present in < 5 samples
  frac_present <- function (x) sum(x > 0) / length(x)
  too_rare <- apply(cts, 1, function (x) frac_present(x) < min_fraction)

  res$tests <- apply(cts[!too_rare,], 1, function (x) kruskal.test(x,s[[var]]))

  medians <- t(apply(props[!too_rare,], 1, tapply, s[[var]], median))
  colnames(medians) <- paste("Median", colnames(medians))
  
  res$df <- data.frame(Taxon = names(res$tests))
  res$df <- cbind(
    res$df, 
    medians,
    Pval = sapply(res$tests, `[[`, "p.value"))
  res$df <- within(res$df, {
    FDR <- p.adjust(Pval, method="fdr")
  })
  res$df <- arrange(res$df, Pval)
  
  plotdf <- melt(props, varnames=c("Taxon", "SampleID"), value.name="Proportion")
  plotdf <- merge(plotdf, res$df, by="Taxon", all.x=T, all.y=F)
  plotdf <- subset(plotdf, FDR < 0.05)
  plotdf <- merge(plotdf, s, by="SampleID", all.x=T, all.y=F)
  plotdf$Taxon <- reorder(plotdf$Taxon, plotdf$Pval)

  res$plot <- ggplot(plotdf) + 
    geom_boxplot(aes_string(x=var, y="Proportion"), coef=200) + theme(text = element_text(size=10),
        axis.text.x = element_text(angle=90, vjust=1)) +
    facet_wrap(~ Taxon, scales="free_y", ncol=2) 
    #theme_classic()

  class(res) <- "taxon_abundance_tests"
  res
}
print.taxon_abundance_tests <- function (x) {
  print(x$plot)
  print(subset(x$df, Pval < 0.05))
}
```

```{r StatEllipse, echo=FALSE}
require(proto)

StatEllipse <- proto(ggplot2:::Stat,
  {
    required_aes <- c("x", "y")
    default_geom <- function(.) GeomPath
  	objname <- "ellipse"

		calculate_groups <- function(., data, scales, ...){
			.super$calculate_groups(., data, scales,...)
		}
		calculate <- function(., data, scales, level = 0.75, segments = 51,...){
      dfn <- 2
      dfd <- length(data$x) - 1
      if (dfd < 3){
      	ellipse <- rbind(c(NA,NA))	
      } else {
          require(MASS)
          v <- cov.trob(cbind(data$x, data$y))
          shape <- v$cov
          center <- v$center
          radius <- sqrt(dfn * qf(level, dfn, dfd))
          angles <- (0:segments) * 2 * pi/segments
          unit.circle <- cbind(cos(angles), sin(angles))
          ellipse <- t(center + radius * t(unit.circle %*% chol(shape)))
      }
    
      ellipse <- as.data.frame(ellipse)
      colnames(ellipse) <- c("x","y")
      return(ellipse)
		}
	}
)

stat_ellipse <- function(mapping=NULL, data=NULL, geom="path", position="identity", ...) {
  StatEllipse$new(mapping=mapping, data=data, geom=geom, position=position, ...)
}
```

## Beta diversity analyses

### Unweighted Unifrac plots, colored by study group

#### All 3 groups at study enrollment (t1)

```{r BetaDiversity_unweighted_overall_plot_t1, fig.height=3, fig.width=4.5, echo=FALSE}
# Make SubjectID a categorical variable
s$SubjectID <- as.factor(s$SubjectID)

s_t1 <- subset(s, Age_bin %in% "1")

unw_t1 <- pcoa_1way(s_t1, unw, var="StudyGroup")
unw_t1$plot + scale_color_manual(values=c("BREASTFED"="red","FORMULA A"="blue","FORMULA B"="green"))
#theme(legend.direction="horizontal", legend.position="bottom")
```

#### Formula B vs. Breastfed at final timepoint (t4)
(no Formula A samples)

```{r BetaDiversity_Weighted_BreastfedFormB_plot_t4, fig.height=3, fig.width=4.5, echo=FALSE}
sBBr_t4 <- subset(s_t4,!(StudyGroup %in% "FORMULA A"))
wtu_sBBr_t4 <- pcoa_1way(sBBr_t4, wtu, var="StudyGroup")
wtu_sBBr_t4$plot + scale_color_manual(values=c("BREASTFED"="red","FORMULA A"="blue","FORMULA B"="green"))
wdf_sBBr_t4 <- droplevels(wtu_sBBr_t4$df)
adonis(wtu_sBBr_t4$distances ~ StudyGroup, data=wdf_sBBr_t4)
```
## --------------------------------------------

#### Unweighted Unifrac, by subject

```{r BetaDiversity_unweightedplot_bysubject, fig.height=3, fig.width=4.5, echo=FALSE}
s$Age_bin <- as.factor(s$Age_bin)

pcoa_by_subject <- function(current_subjectID) {
  s_Br10 <- subset(s, SubjectID %in% current_subjectID)
  unw_Br10 <- pcoa_1way(s_Br10, unw, var="Age_bin")
  unw_Br10$plot + scale_color_manual(values=c("1"="red","2"="orange","3"="green","4"="blue")) + labs(title=current_subjectID)
}

lapply(as.character(unique(s$SubjectID)), pcoa_by_subject)
```


```{r DifferingTaxonAbundance_figure, fig.height=9, echo=FALSE}
test_taxon_abundance(s, cts, var="StudyGroup", a=a )
```

